---
title: "SMART Weight Loss: Analytics for February 2021 Submission"
author: |
    |
date: '`r format(Sys.Date(), "%B %d, %Y")`'
geometry: margin=0.7in
output: 
  pdf_document:
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(fig.pos = 'H')
```


```{r, echo = FALSE, message = FALSE}
library(rmarkdown)
library(knitr)
library(kableExtra)
library(dplyr)
```

# Simply using a training set and a validation set

```{r, echo = FALSE, message = FALSE}
myseed <- 342386787
```

## How many individuals did we consider and what percentage had missing data?
```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
source("month-six-00.R")
source("month-six-01.R")

missdf %>%
  select(tot_rows, percent_day1, percent_day4, percent_day7, miss_analysis) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',8),
        col.names = c("Total No. Individuals","Day 1","Day 4","Day 7","Day 1, 4, or 7"),
        row.names = FALSE,
        escape = FALSE,
        caption = "First Column: Total number of individuals in APP ONLY arm (excluding one individual with data issue); Remaining Columns: Percent with no weights") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```

## Can we identify APP ONLY arm non-responders during the first week of treatment? (`non-response' as defined in previous SMART study at weeks 2, 4, 8)

```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")

tbl_cart_only %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "CART Only") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule1"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule2"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "Overall Decision Rule, Option 2: Non-responder if `day 1' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```



```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule3"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "Overall Decision Rule, Option 3: Non-responder if `day 4 or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule4"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "Overall Decision Rule, Option 4: Non-responder if either `day 1 and day 4' or `day 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```



```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule5"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "Overall Decision Rule, Option 5: Non-responder if at least 2 days (any day) out of the first 7 days is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


## Among individuals in the APP ONLY arm, does non-response status based on overall decision rule (Option 1) predict weight loss success at Month X?


```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule1"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

n_tot <- nrow(dat_reg)
n_participants2 <- nrow(na.omit(dat_reg2))
n_participants3 <- nrow(na.omit(dat_reg3))

tbl_mod_outcome2 %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',4),
        col.names = c("Estimate", "Std. Error", "z value", "Pr(>|z|)"),
        row.names = TRUE,
        escape = FALSE,
        caption = paste("Weight loss success at Month 3: Estimates of Parameters in a Logistic Regression Model"," (N=",n_participants2," participants were used to estimate logistic regression model; ", n_tot-n_participants2," participants had missing weight at 3-mo)", sep="")) %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))

tbl_mod_outcome3 %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',4),
        col.names = c("Estimate", "Std. Error", "z value", "Pr(>|z|)"),
        row.names = TRUE,
        escape = FALSE,
        caption = paste("Weight loss success at Month 6: Estimates of Parameters in a Logistic Regression Model"," (N=",n_participants3," participants were used to estimate logistic regression model; ", n_tot-n_participants3," participants had missing weight at 6-mo)", sep="")) %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```

```{r, echo = FALSE, message = FALSE}
expit <- function(x){
  exp(x)/(1+exp(x))
}
```


```{r, echo = FALSE, message = FALSE}
logodds_success <- tbl_mod_outcome2[1,1] + tbl_mod_outcome2[2,1]
logodds_failure <- tbl_mod_outcome2[1,1]

prob_success <- expit(logodds_success)
prob_failure <- expit(logodds_failure)
odds_success <- prob_success/(1+prob_success)
odds_failure <- prob_failure/(1+prob_failure)
OR <- odds_success/odds_failure

OR_3mo <- round(OR, digits=3)
```

```{r, echo = FALSE, message = FALSE}
logodds_success <- tbl_mod_outcome3[1,1] + tbl_mod_outcome3[2,1]
logodds_failure <- tbl_mod_outcome3[1,1]

prob_success <- expit(logodds_success)
prob_failure <- expit(logodds_failure)
odds_success <- prob_success/(1+prob_success)
odds_failure <- prob_failure/(1+prob_failure)
OR <- odds_success/odds_failure

OR_6mo <- round(OR, digits=3)
```

Odds ratio with respect to these two models is `r OR_3mo` and `r OR_6mo`, respectively.

## Among individuals in the APP ONLY arm, does non-response status based on overall decision rule (Option 5) predict weight loss success at Month X?


```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule5"
source("month-six-00.R")
source("month-six-01.R")
source("month-six-02.R")
source("month-six-03.R")

n_tot <- nrow(dat_reg)
n_participants2 <- nrow(na.omit(dat_reg2))
n_participants3 <- nrow(na.omit(dat_reg3))

tbl_mod_outcome2 %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',4),
        col.names = c("Estimate", "Std. Error", "z value", "Pr(>|z|)"),
        row.names = TRUE,
        escape = FALSE,
        caption = paste("Weight loss success at Month 3: Estimates of Parameters in a Logistic Regression Model"," (N=",n_participants2," participants were used to estimate logistic regression model; ", n_tot-n_participants2," participants had missing weight at 3-mo)", sep="")) %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))

tbl_mod_outcome3 %>%
  round(., digits = 3) %>%
  kable(booktabs = TRUE, align = rep('c',4),
        col.names = c("Estimate", "Std. Error", "z value", "Pr(>|z|)"),
        row.names = TRUE,
        escape = FALSE,
        caption = paste("Weight loss success at Month 6: Estimates of Parameters in a Logistic Regression Model"," (N=",n_participants3," participants were used to estimate logistic regression model; ", n_tot-n_participants3," participants had missing weight at 6-mo)", sep="")) %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```

```{r, echo = FALSE, message = FALSE}
logodds_success <- tbl_mod_outcome2[1,1] + tbl_mod_outcome2[2,1]
logodds_failure <- tbl_mod_outcome2[1,1]

prob_success <- expit(logodds_success)
prob_failure <- expit(logodds_failure)
odds_success <- prob_success/(1+prob_success)
odds_failure <- prob_failure/(1+prob_failure)
OR <- odds_success/odds_failure

OR_3mo <- round(OR, digits=3)
```

```{r, echo = FALSE, message = FALSE}
logodds_success <- tbl_mod_outcome3[1,1] + tbl_mod_outcome3[2,1]
logodds_failure <- tbl_mod_outcome3[1,1]

prob_success <- expit(logodds_success)
prob_failure <- expit(logodds_failure)
odds_success <- prob_success/(1+prob_success)
odds_failure <- prob_failure/(1+prob_failure)
OR <- odds_success/odds_failure

OR_6mo <- round(OR, digits=3)
```

Odds ratio with respect to these two models is `r OR_3mo` and `r OR_6mo`, respectively.

# Use five-fold cross validation


```{r, echo = FALSE, message = FALSE}
use_this_outcome <- "outcome1"
use_miss_rule <- "miss_rule1"
source("month-six-00.R")
```

```{r, echo = FALSE, message = FALSE}
myseed <- 8923412
set.seed(myseed)

cvgroups <- rmultinom(n=nrow(dat), size=1, prob = c(.2, .2, .2, .2, .2))
cvgroups <- t(cvgroups)
cvgroups <- data.frame(cvgroups)
colnames(cvgroups) <- c("cv1","cv2","cv3","cv4","cv5")

# shuffle data before assigning to cv group
shuffleddat <- dat[sample.int(nrow(dat)),]
usedat <- cbind(shuffleddat, cvgroups)

n_cv <- usedat %>% summarise(n_cv1 = sum(cv1), n_cv2 = sum(cv2), n_cv3 = sum(cv3), n_cv4 = sum(cv4), n_cv5 = sum(cv5))

n_cv %>%
  kable(booktabs = TRUE, align = rep('c',5),
        col.names = c("1st Fold", "2nd Fold", "3rd Fold", "4th Fold", "5th Fold"),
        row.names = FALSE,
        escape = FALSE,
        caption = "Total Number of paticipants in each fold") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```



```{r, echo = FALSE, message = FALSE}
# Finally, take subset of individuals who will be used as training data and test data
dat_train <- usedat %>% filter(cv1!=1) %>% ungroup(.)
dat_test <- usedat %>% filter(cv1==1) %>% ungroup(.)

# -----------------------------------------------------------------------------
# What is the outcome that will be used in the classification rule?
# -----------------------------------------------------------------------------
dat_train[["Y"]] <- dat_train[[use_this_outcome]]  # see top of script
dat_test[["Y"]] <- dat_test[[use_this_outcome]]  # see top of script


if(FALSE){
  dat_train <- dat_train %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
  dat_test <- dat_test %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
}


source("month-six-02.R")
source("month-six-03.R")

tbl_cv1 <- tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3) 

tbl_cv1_formatted <- tbl_cv1 %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "1st training set-validation set split: Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```

```{r, echo = FALSE, message = FALSE}
# Finally, take subset of individuals who will be used as training data and test data
dat_train <- usedat %>% filter(cv2!=1) %>% ungroup(.)
dat_test <- usedat %>% filter(cv2==1) %>% ungroup(.)

# -----------------------------------------------------------------------------
# What is the outcome that will be used in the classification rule?
# -----------------------------------------------------------------------------
dat_train[["Y"]] <- dat_train[[use_this_outcome]]  # see top of script
dat_test[["Y"]] <- dat_test[[use_this_outcome]]  # see top of script


if(FALSE){
  dat_train <- dat_train %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
  dat_test <- dat_test %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
}


source("month-six-02.R")
source("month-six-03.R")

tbl_cv2 <- tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3)

tbl_cv2_formatted <- tbl_cv2 %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "2nd training set-validation set split: Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
# Finally, take subset of individuals who will be used as training data and test data
dat_train <- usedat %>% filter(cv3!=1) %>% ungroup(.)
dat_test <- usedat %>% filter(cv3==1) %>% ungroup(.)

# -----------------------------------------------------------------------------
# What is the outcome that will be used in the classification rule?
# -----------------------------------------------------------------------------
dat_train[["Y"]] <- dat_train[[use_this_outcome]]  # see top of script
dat_test[["Y"]] <- dat_test[[use_this_outcome]]  # see top of script


if(FALSE){
  dat_train <- dat_train %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
  dat_test <- dat_test %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
}


source("month-six-02.R")
source("month-six-03.R")

tbl_cv3 <- tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3)

tbl_cv3_formatted <- tbl_cv3 %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "3rd training set-validation set split: Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
# Finally, take subset of individuals who will be used as training data and test data
dat_train <- usedat %>% filter(cv4!=1) %>% ungroup(.)
dat_test <- usedat %>% filter(cv4==1) %>% ungroup(.)

# -----------------------------------------------------------------------------
# What is the outcome that will be used in the classification rule?
# -----------------------------------------------------------------------------
dat_train[["Y"]] <- dat_train[[use_this_outcome]]  # see top of script
dat_test[["Y"]] <- dat_test[[use_this_outcome]]  # see top of script


if(FALSE){
  dat_train <- dat_train %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
  dat_test <- dat_test %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
}


source("month-six-02.R")
source("month-six-03.R")

tbl_cv4 <- tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3)

tbl_cv4_formatted <- tbl_cv4 %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "4th training set-validation set split: Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
# Finally, take subset of individuals who will be used as training data and test data
dat_train <- usedat %>% filter(cv5!=1) %>% ungroup(.)
dat_test <- usedat %>% filter(cv5==1) %>% ungroup(.)

# -----------------------------------------------------------------------------
# What is the outcome that will be used in the classification rule?
# -----------------------------------------------------------------------------
dat_train[["Y"]] <- dat_train[[use_this_outcome]]  # see top of script
dat_test[["Y"]] <- dat_test[[use_this_outcome]]  # see top of script


if(FALSE){
  dat_train <- dat_train %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
  dat_test <- dat_test %>%
    mutate(delta_1 = if_else(is.na(delta_1), 0, delta_1),
           delta_4 = if_else(is.na(delta_4), 0, delta_4),
           delta_7 = if_else(is.na(delta_7), 0, delta_7))
  
}


source("month-six-02.R")
source("month-six-03.R")

tbl_cv5 <- tbl_overall %>%
  select(sensitivity,specificity) %>%
  round(., digits = 3)

tbl_cv5_formatted <- tbl_cv5 %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "5th training set-validation set split: Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
tbl_average <- (tbl_cv1 + tbl_cv2 + tbl_cv3 + tbl_cv4 + tbl_cv5)/5

tbl_average_formatted <- tbl_average %>%
  kable(booktabs = TRUE, align = rep('c',2),
        col.names = c("Sensitivity","Specificity"),
        row.names = TRUE,
        escape = FALSE,
        caption = "Average across five splits: Overall Decision Rule, Option 1: Non-responder if `day 1, 4, or 7' is missing") %>%
  kable_styling(full_width = FALSE, latex_options = c("HOLD_position"))
```


```{r, echo = FALSE, message = FALSE}
tbl_cv1_formatted
tbl_cv2_formatted
tbl_cv3_formatted
tbl_cv4_formatted
tbl_cv5_formatted
tbl_average_formatted
```



